---
title: "BRFSS Dashboard Workflow"
author: "Mandar"
format: html
editor: visual
---

```{r}
install.packages('arrow')
```

```{r setup, include = FALSE}
library(tidyverse)
library(arrow)
library(plotly)
library(DT)
library(stringr)
```

Time to load the data.

```{r load_data, show_col_types = TRUE}
df <- read_csv(" DS_Project.csv")

slice_sample(df, n = 5)

```

Now that we have uploaded the csv file, let us merge the functions to tackle inconsistencies.

```{r merge_functions}
# Merge inconsistent Response IDs
merge_ResponseID <- function(char_vec) {
  char_vec <- str_replace(char_vec,"RESP025","RESP137")
  char_vec <- str_replace(char_vec,"RESP026","RESP172")
  char_vec <- str_replace(char_vec,"RESP029","RESP141")
  char_vec <- str_replace(char_vec,"RESP230","RESP020")
  char_vec <- str_replace(char_vec,"RESP231","RESP020")
  char_vec <- str_replace(char_vec,"RESP232","RESP020")
  char_vec <- str_replace(char_vec,"RESP196","RESP199")
  char_vec <- str_replace(char_vec,"RESP197","RESP199")
  char_vec <- str_replace(char_vec,"RESP198","RESP199")
  char_vec <- str_replace(char_vec,"RESP199","RESP199")
  char_vec <- str_replace(char_vec,"RESP200","RESP008")
  char_vec <- str_replace(char_vec,"RESP194","RESP005")
  char_vec <- str_replace(char_vec,"RESP195","RESP006")
  return(char_vec)
}

# Match new Response labels to the merged IDs
merge_Response <- function(ResponseID, Response) {
  idx <- str_detect(ResponseID,"RESP137"); Response[idx] <- "Employed"
  idx <- str_detect(ResponseID,"RESP172"); Response[idx] <- "Self-employed"
  idx <- str_detect(ResponseID,"RESP141"); Response[idx] <- "Homemaker"
  idx <- str_detect(ResponseID,"RESP020"); Response[idx] <- "$50,000+"
  idx <- str_detect(ResponseID,"RESP199"); Response[idx] <- "Alaskan/American Native, Asian, Pac. Islander,Other"
  idx <- str_detect(ResponseID,"RESP008"); Response[idx] <- "Multiracial"
  idx <- str_detect(ResponseID,"RESP005"); Response[idx] <- "White"
  idx <- str_detect(ResponseID,"RESP006"); Response[idx] <- "Black"
  return(Response)
}

# Merge inconsistent Breakout IDs
merge_BreakoutID <- function(char_vec) {
  char_vec <- str_replace(char_vec,"INCOME01","INCOME1")
  char_vec <- str_replace(char_vec,"INCOME02","INCOME2")
  char_vec <- str_replace(char_vec,"INCOME03","INCOME3")
  char_vec <- str_replace(char_vec,"INCOME04","INCOME4")
  char_vec <- str_replace(char_vec,"INCOME05","INCOME5")
  char_vec <- str_replace(char_vec,"INCOME06","INCOME5")
  char_vec <- str_replace(char_vec,"INCOME07","INCOME5")
  char_vec <- str_replace(char_vec,"RACE01","RACE1")
  char_vec <- str_replace(char_vec,"RACE02","RACE2")
  char_vec <- str_replace(char_vec,"RACE08","RACE3")
  char_vec <- str_replace(char_vec,"RACE04","RACE4")
  char_vec <- str_replace(char_vec,"RACE05","RACE4")
  char_vec <- str_replace(char_vec,"RACE06","RACE4")
  char_vec <- str_replace(char_vec,"RACE03","RACE4")
  char_vec <- str_replace(char_vec,"RACE07","RACE5")
  return(char_vec)
}

# Assign readable names to merged Breakouts
merge_Break_Out <- function(BreakoutID, Break_Out) {
  idx <- str_detect(BreakoutID,"INCOME5"); Break_Out[idx] <- "$50,000+"
  idx <- str_detect(BreakoutID,"RACE1"); Break_Out[idx] <- "White"
  idx <- str_detect(BreakoutID,"RACE2"); Break_Out[idx] <- "Black"
  idx <- str_detect(BreakoutID,"RACE3"); Break_Out[idx] <- "Hispanic"
  idx <- str_detect(BreakoutID,"RACE4"); Break_Out[idx] <- "Alaskan/American Native, Asian, Pac. Islander,Other"
  idx <- str_detect(BreakoutID,"RACE5"); Break_Out[idx] <- "Multiracial"
  return(Break_Out)
}
```

Apply these functions to the uploaded data df to get them implemented. First lets verify some information before the implementation.

Below we verify the kind of questions we have and how they are formatted.

```{r}
unique(df$Question[ str_detect(df$Question, "health") ])
```

```{r}
unique(df$Break_Out[1:20])
```

Let us check the number of inputs each value of Break_Out has in the entire dataset.

```{r}
df %>%
  count(Break_Out) %>%
  arrange(desc(n))
```

Let us perform a sample test aggregation on a given question and onbserve the answers that follow it.

```{r test_aggregation}

question <- "How is your general health?"

df_overall <- df |>
  mutate(
    Data_value = as.numeric(Data_value),
    Sample_Size = as.numeric(Sample_Size)
  ) |>
  filter(
    str_detect(Question, "general health"),
    Break_Out == "Overall",
    !is.na(Data_value),
    !is.na(Sample_Size)
  ) |>
  rename(persons = Sample_Size) |>
  mutate(Sample_Size = persons * 100 / Data_value) |>
  group_by(Response) |>
  summarise(
    agg_ss = sum(Sample_Size, na.rm = TRUE),
    agg_persons = sum(persons, na.rm = TRUE),
    agg_percent = agg_persons * 100 / agg_ss,
    agg_percent_sdev = sqrt(agg_percent * (100 - agg_percent) / agg_ss),
    low_CI = agg_percent - 2 * agg_percent_sdev,
    high_CI = agg_percent + 2 * agg_percent_sdev
  )

df_overall
```

Let us visualise these results using an interactive bar plot.

```{r plot_overall, echo=FALSE}
library(plotly)

plot_ly(
  df_overall,
  x = ~Response,
  y = ~agg_percent,
  type = "bar",
  error_y = list(
    type = "data",
    array = df_overall$high_CI - df_overall$agg_percent,
    arrayminus = df_overall$agg_percent - df_overall$low_CI,
    color = '#444',
    thickness = 1.5,
    width = 6
  )
) %>%
  layout(
    title = "Overall Health Status – 95% Confidence Intervals",
    xaxis = list(title = "Response"),
    yaxis = list(title = "Percentage (%)")
  )
```

Now that the code works for a given catogory, we can implement this to work for any question and any grouping category(eg, Gender, Age, Income).

```{r}
aggregate_question <- function(df,question_pattern, breakout_value = "Overall"){
  df |> mutate(Data_value = as.numeric(Data_value),
               Sample_Size = as.numeric(Sample_Size)
              ) |>
    filter(
      str_detect(Question, regex(question_pattern, ignore_case = TRUE)),
      Break_Out == breakout_value,
      !is.na(Data_value),
      !is.na(Sample_Size)
    ) |>
    rename(persons = Sample_Size) |>
    mutate(Sample_Size = persons * 100 / Data_value) |>
    group_by(Response) |>
    summarise(
      agg_ss = sum(Sample_Size, na.rm = TRUE),
      agg_Persons = sum(persons, na.rm = TRUE),
      agg_Percent = agg_Persons *100 / agg_ss,
      agg_percent_sdev = sqrt(agg_Percent * (100 - agg_Percent) / agg_ss),
      low_CI = agg_Percent - 2 * agg_percent_sdev,
      high_CI = agg_Percent + 2 * agg_percent_sdev
    ) |>
    arrange(desc(agg_Percent))
}
df_overall <- aggregate_question(df, "general health", "Overall")
df_overall
```

```{r}
df_overall <- aggregate_question(df, "general health", "Overall")
df_gender_male <- aggregate_question(df, "general health", "Male")
df_gender_female <- aggregate_question(df, "general health", "Female")
```

Checking what al categories belong to Break_Out

```{r}
df %>% count(Break_Out_Category, Break_Out)
```

```{r plot_function}
plot_confidence_bar <- function(data, question, breakout_value) {
  plot_ly(
    data,
    x = ~Response,
    y = ~agg_Percent,
    type = "bar",
    error_y = list(
      type = "data",
      array = data$high_CI - data$agg_Percent,
      arrayminus = data$agg_Percent - data$low_CI
    )
  ) %>%
    layout(
      title = paste0(question, " – ", breakout_value),
      xaxis = list(title = "Response"),
      yaxis = list(title = "Percentage (%)")
    )
}

```

```{r}
plot_confidence_bar(df_overall, "How is your general health?", "Overall")
plot_confidence_bar(df_gender_male, "How is your general health?", "Male")
```

```{r}
df %>%
  filter(str_detect(Question, "How is your general health?", ignore_case=TRUE)) %>%
  distinct(Break_Out) %>%
  print(n=5)
```

```{r}

df %>% count(Break_Out) %>% arrange(desc(n))

```

```{r}
df %>%
  filter(
    str_detect(Question, regex("Adults who are current e-cigarette users", ignore_case = TRUE)),
    BreakOutCategoryID == "CAT2"
  ) %>% 
  head()
```
